Conceptual and methodological problems with bias detection and avoidance in natural language processing

### 2021-06-15 thesis/03-bayesian.Rmd
Just saving file
Adds fixes

\todo{Explain where you took the embedding from, cite publications first announcing it.}

@@

>>>>

We first encountered the GoogleNews-vectors-negative300 word embedding model in @Bolukbasi2016Man. They use it for their calculations as mentioned in "https://github.com/tolga-b/debiaswe". We decided to compare the results obtained with the use of Reddit model and the ones that we got by GoogleNews model. The details of GoogleNews model can be found here "https://code.google.com/archive/p/word2vec/". One of the main differences between the models is that Reddit word embeggings have 50 dimensions and GoogleNews word embeddings have 300 dimensions. As the dimension is increasing, the vector can capture much more information but this is not always the case. According to some researchers the amount of dimensions should be chosen by taking into account the type of corpus and its features. In our case both Reddit and GoogleNews models are already trained and ready to use so we do not analyse further the choice of hyperparameters to train them.

@@

- humans in religion got closer (should they?)

- change in Gender is really minor, still zero out of HPDI range, note how different is very similar

- small improvement in race, at the price of moving neutral terms closer

### 2021-06-15 thesis/04-discussion.Rmd
Adds fixes

Some points to discuss:

- humans in religion got closer (should they?)

- change in Gender is really minor, still zero out of HPDI range, note how different is very similar

- small improvement in race, at the price of moving neutral terms closer

>>>>

It is worth diving deeper into some observations that arise while analyzing the results.

### 2021-06-15 thesis/_book/_main.tex
Adds fixes

@@

We will now think of cosine distance as the output variable, and
will build a few bayesian models to compare. First, we just build a baseline model which estimates cosine distance to the attributes separately for each protected word. The underlying idea is that different protected words migh in general have different relations to all the attributes and this relations should be our point of departure.

Here is the intution behind the mathematical Bayesian model involved. Our outcome variable is \texttt{cosine\ difference}, which we take to me normally distributed around the predicted mean for a given protected word (that is, we assume the residuals are normally distributed). The simplest model specification is:

@@

\todo{Explain where you took the embedding from, cite publications first announcing it.}

@@

\todo{comment on this}

Some points to discuss:

@@

>>>>

\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

@@

We will now think of cosine distance as the output variable, and
will build a few Bayesian models to compare. First, we just build a baseline model which estimates cosine distance to the attributes separately for each protected word. The underlying idea is that different protected words might in general have different relations to all the attributes and these relations should be our point of departure.

Here is the intuition behind the mathematical Bayesian model involved. Our outcome variable is \texttt{cosine\ difference}, which we take to me normally distributed around the predicted mean for a given protected word (that is, we assume the residuals are normally distributed). The simplest model specification is:

@@

We first encountered the GoogleNews-vectors-negative300 word embedding model in \protect\hyperlink{ref-Bolukbasi2016Man}{Bolukbasi, Chang, Zou, Saligrama, \& Kalai} (\protect\hyperlink{ref-Bolukbasi2016Man}{2016}). They use it for their calculations as mentioned in \url{https://github.com/tolga-b/debiaswe}. We decided to compare the results obtained with the use of Reddit model and the ones that we got by GoogleNews model. The details of GoogleNews model can be found here \url{https://code.google.com/archive/p/word2vec/}. One of the main differences between the models is that Reddit word embeggings have 50 dimensions and GoogleNews word embeddings have 300 dimensions. As the dimension is increasing, the vector can capture much more information but this is not always the case. According to some researchers the amount of dimensions should be chosen by taking into account the type of corpus and its features. In our case both Reddit and GoogleNews models are already trained and ready to use so we do not analyse further the choice of hyperparameters to train them.

@@

@@

It is worth diving deeper into some observations that arise while analyzing the results.

\begin{itemize}
\tightlist
\item
  89\% highest posterior density intervals
\end{itemize}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-Bolukbasi2016Man}{}%
Bolukbasi, T., Chang, K.-W., Zou, J. Y., Saligrama, V., \& Kalai, A. (2016). Man is to computer programmer as woman is to homemaker? Debiasing word embeddings. \emph{CoRR}, \emph{abs/1607.06520}. Retrieved from \url{http://arxiv.org/abs/1607.06520}

\end{CSLReferences}
